---
title: 'Exercise 2. Data analysis'
author: 'Sasha D. Hafner'
date: "`r format(Sys.time(), '%d %B, %Y')`"
urlcolor: blue
---

**R data.table solution**

# Overview
In this exercise you will analyze emission of methane from pig slurry samples incubated in a laboratory experiment.
Multiple bottles were used as reactors in a completely crossed factorial experiment with 2 temperatures and 2 headspace gases.
Measurement data were kindly provided by Frederik Dalby.

### My suggestion
I recommend using response feature analysis with calculated total emission as the response variable, fitting a linear model using `lm()` in R or `ols()` in Python.
But you could use other responses.

# 1. Read and check data
Read in the data in the file `slurry_emis.csv`.
Check the data.
The relevant columns are 

* `reactor`: bottle key
* `ch4`: methane concentration in bottle exhaust in ppmv
* `flow`: rate of gas flow through the bottle in L/min
* `day`: time of measurement from setup in d
* `gas`: headspace gas
* `temp`: incubation temperature

## Load packages

```{r}
library(data.table)
library(rmarkdown)
library(ggplot2)
```

## Function
```{r}
source('../functions-R/dfsumm.R')
```

## Load data

```{r}
dat <- fread('../data/slurry_emis.csv')
```

## Check data

```{r}
dat
```

```{r}
summary(dat)
```

```{r}
dfsumm(dat)
```

```{r}
dim(dat)
```

```{r}
table(dat[, reactor])
```

# 2. Data analysis
*Use an appropriate approach to determine if the data show that incubation temperature and headspace gas affect methane emission.
Quantify any effect.*

## Calculate methane flow rate
First we need mass-based methane concentration.
From ideal gas law `pv = nRT` we get `Mn / v = p / RT`.
Assume 20 degress C and 1 atm (would be good to check if these were really our data).
Temperature should be at the point of flow rate measurement, right?
And `gas_constant = 8.2057E-5` for the universal gas constant in atm m^3^k mol^-1^ K^-1^.
So first, concentration in g/m3:

```{r}
dat[, cmch4 := ch4 * 1.0 / 1E6 / 8.2057E-5 / (273.15 + 20)]
dat
```

And flow in g/d:

```{r}
dat[, qch4 := cmch4 * flow / 1000 * 1400]
dat
```

## Get integration function

```{r}
source('../functions-R/mintegrate.R')
```

## Integrate

```{r}
dim(dat)
dat <- dat[!is.na(qch4), ]
dim(dat)

dat[, ech4 := mintegrate(day, qch4), by = .(reactor, gas, temp)]

dattot <- dat[, .(ech4 = mintegrate(day, qch4, value = 'total')), by = .(reactor, gas, temp)]
```

## Plot

```{r}
ggplot(dattot, aes(gas, ech4, colour = factor(temp))) +
  geom_jitter(height = 0, width = 0.2)
```

```{r}
ggplot(dat, aes(day, qch4, colour = factor(temp), lty = factor(gas), group = reactor)) +
  geom_line()
```

```{r}
ggplot(dat, aes(day, ech4, colour = factor(temp), lty = factor(gas), group = reactor)) +
  geom_line()
```

## Replication

```{r}
table(dattot[, .(gas, temp)])
table(dattot[, .(reactor, gas, temp)])
```

Drop background.

```{r}
dattot <- dattot[!reactor == 'bg', ]
```

```{r}
table(dattot[, .(gas, temp)])
```

```{r}
ggplot(dattot, aes(gas, ech4, colour = factor(temp))) +
  geom_jitter(height = 0, width = 0.2)
```

```{r}
ggplot(dattot, aes(gas, ech4, fill = factor(temp))) +
  geom_boxplot()
```

Look at log transformation.
Looks like constant variance assumption would be better.
We can compare with and without.

```{r}
ggplot(dattot, aes(gas, log10(ech4), fill = factor(temp))) +
  geom_boxplot()
```

## Statistical comparisons

```{r}
dattot[, temp := factor(temp)]
dattot[, gas := factor(gas)]
```

```{r}
mod1 <- aov(ech4 ~ temp*gas, data = dattot)
summary(mod1)
```

```{r}
TukeyHSD(mod1)
```

```{r}
plot(mod1, ask = FALSE)
```

```{r}
confint(mod1)
```

Now with transformation.

```{r}
mod2 <- aov(log10(ech4) ~ temp*gas, data = dattot)
summary(mod2)
```

```{r}
TukeyHSD(mod2)
```

```{r}
plot(mod2, ask = FALSE)
```

```{r}
confint(mod2)
```

Back-transform.

```{r}
10^confint(mod2) - 1
```
